\chapter[\ciovzerotitleemph{}]{\ciovzerolongtitleemph{}}
\chaptermark{\ciovzeroshorttitlenoemph{}}
\label{ciov0}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview of the Library}
\label{ciov0:sdpr0}

%LaTeX won't substitute into \index{} below, so had to use the expanded
%name.
\index{LibNum@\emph{LibNum}}The library described in this document,
\emph{\productbasenameshort{} Version \productversion{}} (or more compactly,
\emph{\productbasenameshort{}-\productversion{}} or
\emph{\productbasenameshort{}}) is a microcontroller-centric
multi-threaded C/C++-callable 
arith\-me\-tic/\-math\-e\-mat\-i\-cal/\-cryp\-to\-graph\-ic 
software library.

The library includes functionality for: 

\begin{itemize}
\item \emph{Arithmetic:}
      fundamental\footnote{\emph{Fundamental integer} is used here
      to mean an integer type
      specified by the C standard and built into the compiler.} integer,
      large\footnote{\emph{Large integer} is used here to mean an integer
      larger than any fundamental integer type.  Such integers are implemented as
      arrays of fundamental integers.} integer, fixed-point, large fixed-point,
      and float\-ing-point addition, subtraction,
      multiplication, and division.
\item \emph{Standard functions:}
      root extraction, logarithm and exponentiation, trigonometric, and similar
      functions.
\item \emph{Mathematical algorithms:}
      fundamental integer, large integer, fixed-point, large fixed-point,
      floating-point, and vector algorithms.
\item \emph{Control system elements and linear filtering:}
      low/high-pass filtering, scaling, differentiation,
      and integration.
\item \emph{Non-linear filtering:}
      discrete input debouncing, and vertical counters.
\item \emph{Non-cryptographic hashes:}
      checksums, CRC's.
\item \emph{Cryptographic hashes:}
      hashes from the SHA-2 family (SHA-224, SHA-256,
      SHA-384, SHA-512, SHA-512/224, and SHA-512/256).
\item \emph{Cryptography:}
      ciphers and PRNGs.
\item \emph{Searching, Sorting, Insertion, and Deletion:}
      searching for elements in unsorted and sorted tables and arrays,
      sorting tables and arrays, inserting and deleting elements from
      unsorted and sorted tables and arrays.
\item \emph{Bit Manipulation:}
      bit cardinality, searches for set and clear bits, bit manipulation.
\item \emph{Utility:}
      block memory
      operations and array operations.
\end{itemize}

\emph{\productbasenameshort{}} has the following features that make it suitable
for use with microcontrollers in embedded systems.

\begin{itemize}
\item The library contains a rich set of functions that work with fixed-size operands.
      (Functions that work with fixed-size operands are often preferred in
      embedded systems because they execute more quickly and often have a smaller
      FLASH footprint.)
\item The library is designed for minimal FLASH footprint.
      (The library uses generation tools that can provide only the source code
      necessary to resolve linker references.  This gives minimal
      FLASH footprint and better compatibility with debuggers,
      as debuggers often work better with source code incorporated into
      a project than with proper libaries.)
\item The library functions are thread-safe, so they can be used in
      multi-threaded applications.
\item The low-level operations of the library are implemented in assembly language
      for efficiency.  (Popular platforms are supported, and it is easy
      to add assembly language for additional platforms. There is a C-language fallback
      for unsupported platforms.)
\item The library functions use minimal stack space.  (The RAM utilization
      paradigm of the library is carefully designed to minimize stack
      consumption.)
\end{itemize}

The features that make \emph{LibNum} suitable for embedded systems do
not make it unsuitable for personal computers.  The differences in likely
utilization of \emph{LibNum} between embedded systems and personal
computer applications are likely:

\begin{itemize}
\item Embedded systems are more likely to use the functions of
      \emph{LibNum} that are useful in control applications
      (debouncing, filtering, etc.).
\item Embedded systems are more likely to use the functions of
      \emph{LibNum} that accept fixed-size operands, whereas
      personal computer applications are more likely to use
      the functions that accept operands of arbitrary size.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
\label{ciov0:sack0}
 
I'm very grateful to the authors of the GMP. Many concepts
of \emph{LibNum} were borrowed from the GMP.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Feedback, Suggestions, and Collaboration Opportunities}
\label{ciov0:sfbk0}

I welcome all feedback and suggestions about the library and 
documentation, and I welcome all opportunities to collaborate to extend 
the functionality or supported platforms of the library.  

Please feel free to correspond with me at \texttt{dashley@\-gmail.com}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Licensing and License Interpretation}
\label{ciov0:slip0}

The \emph{\productbasenameshort{}} library and all 
documentation is licensed under
\index{Unlicense@\emph{The Unlicense}}\emph{The Unlicense}.  The 
license text below (\emph{The Unlicense}) is included in the 
\emph{LICENSE} file in the root directory of the project.  

\begin{small}
\begin{verbatim}
This is free and unencumbered software released into the public domain.

Anyone is free to copy, modify, publish, use, compile, sell, or
distribute this software, either in source code form or as a compiled
binary, for any purpose, commercial or non-commercial, and by any
means.

In jurisdictions that recognize copyright laws, the author or authors
of this software dedicate any and all copyright interest in the
software to the public domain. We make this dedication for the benefit
of the public at large and to the detriment of our heirs and
successors. We intend this dedication to be an overt act of
relinquishment in perpetuity of all present and future rights to this
software under copyright law.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND,
EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
IN NO EVENT SHALL THE AUTHORS BE LIABLE FOR ANY CLAIM, DAMAGES OR
OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE,
ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR
OTHER DEALINGS IN THE SOFTWARE.

For more information, please refer to <https://unlicense.org>
\end{verbatim}
\end{small}

My interpretation of \emph{The Unlicense} is that it places no restrictions
on a user of the software, other than that they may not litigate against
the authors or curators of the software.  In particular, the software may
be used in any product, including embedded products, without the obligation
to notify the end user or to publish changes.

All documentation and images, in any form (source, intermediate, final, .PDF,
.DVI, etc.), including this document in any form, are also licensed under [the same
terms as] \emph{The Unlicense}. (The documentation is closely tied to the software,
and any errors in the software are likely to be mirrored in the documentation,
and vice-versa.)  With respect to all documentation in any form, including this
document in any form, per the
terms of \emph{The Unlicense}, there is no warranty of any kind, no liability,
no obligation to notify the end user, and
no obligation to publish changes.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Detailed Description of this Document}
\label{ciov0:sotd0}

\emph{\textbf{Part \ref{pgeninfo}: \pgeninfolongtitle{}}} provides
general information about \emph{\productbasenameshort{}}.

\begin{itemize}
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{ciov0}: 
      \ciovzerolongtitlenoemph{}}} provides general information
      about \emph{\productbasenameshort{}}.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cldd0}: 
      \clddzerolongtitlenoemph{}}}
      provides a detailed description 
      of the features, characteristics, and design rules of
      of \emph{\productbasenameshort{}}\@.  The specific
      functions and constant data structures 
      in \emph{\productbasenameshort{}} are described in
      Chapters \ref{cuuc0} through \ref{ccip1}.
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{clgn0}: 
      \clgnzerolongtitlenoemph{}}} documents the
      \emph{\productbasenameshort{}} generator, the program that
      packages \emph{\productbasenameshort{}} in multiple forms and
      can customize it for minimal footprint.
\end{itemize}

\emph{\textbf{Part \ref{ptechbackground}: \ptechbackgroundlongtitle{}}}
provides mathematical and technical background information useful in
understanding the library or applying it effectively.

\begin{itemize}
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cgct0}: 
      \cgctzerolongtitlenoemph{}}} provides technical background on a variety of general 
      computing topics.
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cbal0}: 
      \cbalzerolongtitlenoemph{}}} provides technical background on 
      Boolean algebra and its application to digital logic, simplifying software 
      constructs, and vertical counters.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cpta0}: 
      \cptazerolongtitlenoemph{}}} provides technical background 
      related to the design of state machines and stateful systems.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cnth0}: 
      \cnthzerolongtitlenoemph{}}} provides technical background on number theory.  (Number 
      theory is the branch of mathematics that deals with positive integers and 
      their properties, and is the basis of much of modern cryptography.) 
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{crla1}: 
      \crlaonelongtitlenoemph{}}} provides technical background and algorithms 
      surrounding approximating a real number with a rational number (for 
      example, 355/113 is the best approximation to $\pi$ with numerator and 
      denominator not exceeding 1,000).  In some scenarios, using a rational 
      approximation to a real number is the most efficient way to perform a 
      calculation.  Efficiently finding the best rational approximation to a 
      real number requires\footnote{For example, suppose it is necessary to find 
      the best rational approximation to $\pi$ with numerator and denominator 
      not exceeding $2^{256}-1$.  No $O(n)$ algorithm will suffice, as 
      $2^{256}-1$ is too large.  The actual best approximation is 
      110,\-859,\-348,\-167,\-216,\-110,\-469,\-450,\-% 
      301,\-947,\-424,\-979,\-270,\-176,\-705,\-044,\-% 
      426,\-927,\-758,\-107,\-559,\-970,\-506,\-471,\-% 
      518,\-580 / 35,\-287,\-626,\-497,\-515,\-783,\-% 
      106,\-785,\-958,\-721,\-190,\-659,\-891,\-744,\-% 
      246,\-448,\-011,\-713,\-951,\-598,\-935,\-229,\-% 
      278,\-291,\-942,\-269---a result that can only be obtained using the better 
      algorithm presented in this chapter.} results from number theory, which 
      are presented in this chapter.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{ccth0}: 
      \ccthzerolongtitlenoemph{}}} provides technical background on error detecting codes, 
      error correcting codes, and other topics from coding theory.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cnna0}: 
      \cnnazerolongtitlenoemph{}}} provides technical background on a variety 
      non-numerical algorithms that appear in the library, such as algorithms 
      for searching and sorting.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{caal0}: 
      \caalzerolongtitlenoemph{}}} provides technical 
      background on integer arithmetic algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cmal0}: 
      \cmalzerolongtitlenoemph{}}} provides technical 
      background on integer mathematical algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{caal2}: 
      \caaltwolongtitlenoemph{}}} provides technical 
      background on fixed-point arithmetic algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cmal2}: 
      \cmaltwolongtitlenoemph{}}} provides 
      technical background on fixed-point mathematical algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{craa0}: 
      \craazerolongtitlenoemph{}}} 
      provides technical background on real and floating-point arithmetic 
      algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{crma0}: 
      \crmazerolongtitlenoemph{}}} 
      provides technical background on real and floating-point mathematical 
      algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{clfc0}: 
      \clfczerolongtitlenoemph{}}} provides technical background 
      on linear filters and control system elements.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cnlf0}: 
      \cnlfzerolongtitlenoemph{}}} provides technical background on 
      non-linear filters and debouncing.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{crng0}: 
      \crngzerolongtitlenoemph{}}} provides technical background 
      on the generation of random and pseudo-random numbers.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cnch0}: 
      \cnchzerolongtitlenoemph{}}} provides technical background on 
      non-cryptographic hashes.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cchs0}: 
      \cchszerolongtitlenoemph{}}} provides technical background on cryptographic 
      hashes.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cskc0}: 
      \cskczerolongtitlenoemph{}}} provides technical background on 
      symmetric key ciphers and related algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cakc0}: 
      \cakczerolongtitlenoemph{}}} provides technical background on 
      asymmetric key ciphers and related algorithms.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cmat0}: 
      \cmatzerolongtitlenoemph{}}} provides technical 
      background on mathematical and algorithmic topics that could not easily be 
      categorized elsewhere.  
\end{itemize}

\emph{\textbf{Part \ref{plibdoc}: \plibdoclongtitle{}}} describes
usage of the library and the actual library functions.

\begin{itemize}
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cuuc0}: 
      \cuuczerolongtitlenoemph{}}} explains how to use 
      \emph{\productbasenameshort{}} in a project.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cnef0}: 
      \cnefzerolongtitlenoemph{}}} documents functions that provide 
      useful non-data-driven functionality.  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cbmf0}: 
      \cbmfzerolongtitlenoemph{}}} documents functions that operate on blocks of 
      memory (setting, copying, shifting, etc.).  
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{csea0}: 
      \cseazerolongtitlenoemph{}}} documents functions that perform searches; such as 
      linear searches and binary searches.
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{csol0}: 
%      Sort Functions}} documents functions that re-order arrays.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cami0}: 
%      Array Manipulation Functions}} documents functions that manipulate arrays.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cbsf0}: 
%      Bit-Mapped Set Functions}} documents functions that operate on sets 
%      represented as arrays of bits.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cvco0}: 
%      Vertical Counter Functions}} documents functions that implement vertical 
%      counters.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cafn0}: 
%      Basic Data Type Integer Utility and Arithmetic Functions}} documents 
%      utility and arithmetic functions that operate on basic (built into the 
%      compiler) character and integer types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cbaf0}: 
%      Basic Data Type Integer Mathematical Functions}} documents mathematical 
%      functions that operate on basic (built into the compiler) character and 
%      integer types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cfpa0}: 
%      Basic Data Type Fixed-Point Utility and Arithmetic Functions}} documents 
%      utility and arithmetic functions that operate on fixed-point values stored 
%      in basic (built into the compiler) character and integer types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cfpa1}: 
%      Basic Data Type Fixed-Point Mathematical Functions}} documents 
%      mathematical functions that operate on fixed-point values stored in basic 
%      (built into the compiler) character and integer types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{caal1}: 
%      Basic Data Type Floating-Point Utility and Arithmetic Functions}} 
%      documents utility and arithmetic functions that operate on basic (built 
%      into the compiler) floating-point types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cafn1}: 
%      Basic Data Type Floating-Point Mathematical Functions}} documents 
%      mathematical functions that operate on basic (built into the compiler) 
%      floating-point types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{claf0}: 
%      Large Integer Utility and Arithmetic Functions}} documents utility and 
%      arithmetic functions that involve large integers.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{claf1}: 
%      Large-Integer Mathematical Functions}} documents mathematical functions 
%      that involve large integers.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cfpa2}: 
%      Large Fixed-Point Utility and Arithmetic Functions}} documents utility and 
%      arithmetic functions that operate on fixed-point values stored as large 
%      integers.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cfpa3}: 
%      Large Fixed-Point Mathematical Functions}} documents mathematical 
%      functions that operate on fixed-point values stored as large integers.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{claf2}: 
%      Large Floating-Point Utility and Arithmetic Functions}} documents utility 
%      and arithmetic functions that perform arithmetic on floating-point types 
%      larger than basic floating-point data types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{claf3}: 
%      Large Floating-Point Mathematical Functions}} documents utility and 
%      arithmetic functions that perform arithmetic on floating-point types 
%      larger than basic floating-point data types.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{clfi0}: 
%      Linear Filter and Control System Component Functions}} documents functions 
%      that implement classic discrete-time linear filters.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cnfi0}: 
%      Non-Linear Filter Functions}} documents functions that implement 
%      non-linear discrete-time filters.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{crng1}: 
%      Pseudo-Random Number Generation Functions}} documents functions that generate
%      pseudo-random numbers by a variety of algorithms.
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{ccrc0}: 
%      Non-Cryptographic Hash Functions}} documents functions that calculate and 
%      manipulate CRCs, checksums, and non-cryptographic hashes.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{ccrh0}: 
%      Cryptographic Hash Functions}} documents functions that implement 
%      cryptographic hashes.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{ccip0}: 
%      Symmetric Cipher Functions}} documents symmetric cipher functions that 
%      encrypt and decrypt data.  
%\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{ccip1}: 
%      Asymmetric Cipher Functions}} documents asymmetric cipher functions that 
%      encrypt and decrypt data.  
\end{itemize}
%
\emph{\textbf{Part \ref{pdevconinfo}: \pdevconinfolongtitle{}}} provides 
information about building the library from source code and
extending the library.

\begin{itemize}
\item \emph{\textbf{Chapter\postchapterwordnonstretchable{}\ref{cbpc0}: 
      \cbpczerolongtitlenoemph{}}} 
      documents how to generate the library source code from
      the templates, how to modify the templates, how to
      modify the generator, how to add new processor variants,
      how to tune the library, how to build the library, and how
      to test the library.  Design and coding standards are also
      documented.
\end{itemize}

\emph{\textbf{Part \ref{pappbibidx}: \pappbibidxlongtitle{}}}
provides glossaries, references, and an index.  Individuals, 
products, companies, websites, and Internet newsgroups are 
cited in the same framework as traditional references in 
order to provide the reader with more resources to obtain 
information.  

\begin{itemize}
\item \emph{\textbf{Appendix\postchapterwordnonstretchable{}\ref{cglo0}: 
      \cglozerolongtitlenoemph{}}} 
      TBD.
\end{itemize}
